Fast and Error-Adaptive Count-Distinct Sketch \\for Influence Maximization

### 2021-01-22 paper.tex
Update on Overleaf.

\IEEEtitleabstractindextext{%
\begin{abstract}
Influence maximization~(IM) is the problem of finding a seed vertex set that maximizes the expected number of vertices influenced under a given diffusion model. Due to the NP-Hardness of finding an optimal seed set, approximation methods are frequently used for IM. In addition to high-quality yet expensive approximation algorithms, to meet with the scale of today's networks, lightweight, sketch-based approaches, which do not simulate the process exactly, have been proposed in the literature. As expected, the quality of the seed sets for the latter is inferior to those of the former. In this work, we describe a fast, error-adaptive approach that leverages Count-Distinct sketches and the recently proposed hash-based, fused sampling. To estimate the vertex set sizes influenced throughout the simulations, we use per-vertex Flajolet-Martin sketches where each sketch register corresponds to a sample sub-graph. During the diffusion simulations, the reach set cardinalities of a single vertex are stored in adjacent registers in memory. This allows the proposed algorithm to find the number of influenced vertices in a single simulation step in different simulations at once. In addition, thanks to its efficiency and scalability of our parallel implementation can rebuild the sketches after observing large estimation errors. Our experimental results show that the proposed method finds high-quality seed sets while being up-to $1.7\times$--$27.9\times$ faster on average than a state-of-the-art approximation algorithm. In addition, it is shown to be $3.5\times$--$11.1\times$ faster than a sketch-based approach while producing seed sets with $3\%$--$12\%$ better influence scores.
\end{abstract}
% Note that keywords are not normally used for peer review papers.
}

@@

The study of efficient information/influence propagation in networks becomes an important area of research with several applications from many fields, such as viral marketing~\cite{leskovec2007dynamics, trusov2009effects}, social media analysis~\cite{zeng2010social, moreno2004dynamics}, and recommendation systems~\cite{lu2012recommender}.
As the study of these networks is imperative for educational, political, economic, and social purposes, a high-quality seed set to initiate the diffusion may have critical importance.
Furthermore, since the diffusion analysis may be time-critical, or increasing its coverage may be expensive in terms of computing resources and time, novel and efficient approaches to find good vertex sets that propagate the information/influence effectively are essential.

Influence maximization is the problem of finding a subset $S \subset V$ of $K$ vertices in a graph $G = (V, E)$ with the vertex set $V$ and edge set $E$ such that $S$ reaches the maximum reachability under some diffusion model $M$. In addition to introducing the problem, Kempe et al.~\cite{kempe2003maximizing} proved it to be NP-hard and provided a greedy Monte-Carlo approach that has a constant approximation for the optimal solution. This greedy approach is one of the most frequently applied algorithms for IM. The time complexity of the greedy algorithm, estimating the $\sigma$ influence score, running $R$ simulations, and selecting $K$ seed vertices is $\mathcal{O}(KRn\sigma)$ for a graph with $n$ vertices. Although they perform well in terms of seed-set quality, in consequence, the greedy Monte-Carlo solutions are impractical for real-life networks featuring millions of vertices. Due to this reason, many heuristics and proxy methods have been proposed in the literature~\cite{MixGreedy, narayanam2010shapley, kimura2007extracting, chen2010PMIA,chen2010LDAG, kim2013scalable, cohen2014sketch, goyal2011simpath, jung2012irie,cheng2014imrank,liu2014influence,galhotra2016holistic}.

@@

In this work;
\begin{itemize}[leftmargin=*]
\item We propose, \acro\footnote{\url{https://github.com/ggokturk/infuser}}, an open-source, a blazing-fast, and accurate Influence Maximization algorithm. The proposed scheme samples the edges as they are traversed across several simulations. Thus, sampling, diffusion, and count-distinct processes are fused for all simulations.

@@

That is when the simulations traverse the same edge within close time intervals, the edge is accessed only once as long as the simulation timelines allow \acro to do so. The suggested approach, therefore, decreases the pressure on the memory management unit. In relation, for the cascade model, we employ a vector compute unit, near maximum efficiency to regularize memory accesses. 
%\item In addition to instruction-level parallelism, we parallelize \acro on multi-core architectures.

\item \acro{} can process very large graphs with millions of vertices and hundreds of millions of edges under a minute, without compromising the quality of results. For large graphs, due to its race condition indifferent nature, performance scales near linearly with the number of threads available. While processing the graph, only we few GBs of memory is used, mostly for storing the graph itself.

\item We evaluate the performance, memory consumption, and influence score with a state-of-the-art influence maximization algorithms {\sc Skim}~\cite{cohen2014sketch}, {\sc Tim+}~\cite{tim}, {\sc Imm}~\cite{minutoli2019fast} in order to accurately position the performance of \acro{} in the IM literature. The experiments show that \acro is $\minspeedup\times$—$\maxspeedup\times$ faster than state-of-the-art while still being better in comparison in terms of scores, and using a comparable amount of memory.

@@

Let $G = (V,E)$ be a directed graph where the $n$ vertices in $V$ represent the agents, and $m$ edges in $E$ represent the relations between the agents in $V$.
The neighborhood of a vertex $u \in V$ is denoted as $\Gamma_{G}(v) = \{v: (u,v) \in E\}$. 
A graph $G' = (V',E')$ is a sub-graph of $G$ if $V' \subseteq V$ and $E' \subseteq E$. 
Diffusion probability on edge $(u, v) \in G$ is noted as $w_{u,v}$.
$w_{u,v}$ can be determined either by the diffusion model or according to the strength of $u$ and $v$'s relationship in the data.

\begin{table}[!ht]
    \caption{Table of notations}
    \label{tab:notation}
    \centering
    \begin{tabular}{|l|p{0.7\linewidth}|}
        \hline
        Variable & Definition  \\
        \hline
        $G = (V,E)$     & Graph $G$ with vertices $V$ and edges $E$ \\
        $\Gamma_G(v)$   & Neighborhood of incoming edges to vertex $v$ in graph $G$\\ %%FIXME
        $w_{u,v}$       & Probability of $u$ directly influencing $v$ \\
        %$SCC(v) $       & Strongly connected component of vertex $v$\\
        $R_{G}(v)$      & Reachability set of vertex $v$ on graph $G$\\
%        $\overline{R_{G}(v)} & Complement of the vertex set $R_{G}(v)$\\
        \hline\hline
        $S$             & Seed set to maximize influence\\
        $K$             & Size of the seed set\\
        $\mathcal{R}$   & Number of Monte-Carlo simulations performed\\
        $\sigma_{G}(S)$ & Influence score of $S$ in $G$, i.e., expected number of vertices reached from $S$ in $G$\\
        % $\sigma_{G}{(S,v)}$          & Marginal influence gain by adding vertex $v$ to seed set $S$\\
\hline\hline
        %$p$             & Sampling probability for edges\\
%        $P(s,v)_r $     & Random probability generated for selecting edge vertices $s$ to $v$ in simulation $r$\\
        $X_{r..n}$      & Slice of Vector X between r and r+n(not including)\\
        $h(u,v)$        & Hash function for edge $\{u,v\}$\\
        $h_{max}$       & Maximum value hash function $h$ can return\\
        %$X_r$           & Random number/hash generated for simulation $r$  \\
        \hline\hline
        $B$             & Batch size\\
        $e$             & Estimated reachability set size\\
        % $[a, \ldots, a]_B$      & Vector of size $B$, contains all $a$\\
        $M_u[j]$        & $j$th sketch register for vertex $u$\\
        $\varsigma $    & Influence gained before last sketch build\\
        $\sigma $       & Influence Score\\
        $\delta$        & Marginal Gain after last sketch build\\
        $\epsilon_{g}$    & Global estimation error threshold\\
        $\epsilon_{l}$    & Local estimation error threshold\\ 
        $\epsilon_{c}$    & Non-convergenced vertex threshold\\
      %  $x[[i, j]]$   & Slice of vector  $x$ between indices $i$ and $j$, not including $j$ \\ 
        \hline         
    \end{tabular}
\end{table}
\subsection{Influence Maximization}

Influence Maximization aims to find a seed set $S \subseteq V$ among all possible size $K$ subsets of $V$ that maximizes an {\em influence spread function} $\sigma$  when the diffusion process is initiated from $S$. Although we focus on undirected graphs, for IM, the graph can be directed or undirected depending on the initial construction. Figure~\ref{fig:xx} shows a~(Fig.~\ref{fig:ic}) and directed~(Fig.~\ref{fig:wc})
graph for which the weights on the edges are diffusion/influence probabilities.

\begin{figure}[!ht] 
    \centering
  \subfloat[\small{IC}\label{fig:ic}]{%
       \includegraphics[width=0.50\linewidth]{images/ic.pdf}}
  \subfloat[\small{WC}\label{fig:wc}]{%
        \includegraphics[width=0.50\linewidth]{images/wc.pdf}}
    \\% IMAGES HERE
  \caption{\protect\subref{fig:ic} 
The directed graph $G = (V, E)$ for Independent Cascade with independent diffusion probabilities. 
\protect\subref{fig:wc}
The directed graph was obtained from the directed one by setting the diffusion probabilities of incoming edges to $1 / |\Gamma_G(v)|$ for each vertex $v \in V$. 
  }
  %\label{fig_ic_model} 
  \label{fig:xx} 
\end{figure}

The $\sigma_{G, M}(\cdot)$ influence spread function calculates the {\em estimated} number of agents / nodes / vertices affected~(activated) through a $M$ diffusion model. 
We eliminate $M$ from the notation for the sake of elegant simplicity; in the remainder of the text, $\sigma_{G}$ refers to $\sigma_{G, M}$.
In the literature, {\em independent} and {\em weighted cascade}~(IC and WC), and 
{\em linear threshold}~(LT)~\cite{kempe2003maximizing} are some of the widely recognized diffusion models for IM.

@@

\item{\bf Linear threshold} generalizes the cascade models and activates the $v$ vertex once the cumulative activation coming from the neighbors of $v$ exceeds the $\theta v$ threshold. 
All the $\{u, v\}$~(or $(u, v)$ in the directed graph) edges 
with active $u$ vertices are taken into account in the process. Vertex $v$ is activated~\cite{kempe2003maximizing} when the number of these edge weights reaches $\theta v$.  
Therefore, the cascade models are special variants of the linear threshold model with $\theta v = 0 $.

\item {\bf Triggering} proposed by Kempe et al.\cite{kempe2003maximizing} to generalize of previous models mentioned above. In the triggering model, each neighbor has a probability to influence vertex $v$. The diffusion process chooses a random subset of vertices called the "triggering set" to activate vertices at each instance.
\end{itemize}

The complexity analysis stays consistent for many diffusion models including {\em Independent Cascade}~(IC), {\em Linear Threshold}~(LT),  models. We concentrate on the independent cascade model in this paper, but the proposed methods are also relevant to the other models in the literature.

\subsection{Count-Distinct Sketch}\label{sec:sketch}
The count-distinct problem is the problem of finding the number of distinct elements in a stream with non-distinct elements. Computing reachability sets of a vertex is a similar problem; the cardinality of all vertices visited is calculated while traversing sample sub-graphs from the given vertex. Exact cardinality calculations require memory proportional to the cardinality.

@@

The methods based on Reverse reachability and Bottom-Up Traversal utilizes this property directly to merge reachability sets of connected vertices to estimate vertices influence.
MixGreedy method goes one step further; it utilizes that all vertices in a connected component that have the same reachability set in undirected graphs. So, for a sample graph, all reachability sets can be found in a single graph traversal.

For directed graphs, storing reachability sets for all vertices and merging these sets are infeasible for non-trivial graphs. 
If one-hot vectors are used to store reachability sets for constant insertion time, $O(n^2\mathcal{R})$ bits of memory is required and each merge operation has $O(n)$ time complexity. 
if Disjoints sets are used for storing reachability sets; $(n\bar{\sigma}\mathcal{R})$ memory is required to store all reachability sets, and each merge operation has $O(Ackermann'(\sigma))$ complexity.

@@

Flajolet–Martin algorithm only stores how rare elements are in a multi-set.The rarity of the elements is often computed by counting leading zeros in the item's hash values.  The registers are initialized with zero values in the beginning. Items are added by storing the longest leading zero count. Cardinality estimation can be done by taking 2 to the power of the stored value. Multiple registers and hash functions, $M[j]$ and $h_j$, are commonly used to reduce variance.
Adding item $x$ to sketch $M$, and its result $M'$ is shown in equation~\ref{eq:sketch-add}.

@@

The merge operation for two sketches, $M_u$,$M_v$, can be performed by computing the pairwise maximum of the registers. Formally; 
\begin{equation}
    \label{eq:sketch-merge}
    M'[j] = max(M_u[j],M_v[j]) ~ \forall j\in(0,J]
\end{equation} 
where $M'$ is union of set and $J$ is number of sketch registers.

Average of longest running zeros can be used compute the cardinality and the result is divided to correction factor $\phi \approx 0.77351$ to correct hash collisions.
\begin{equation}
    \label{eq:sketch-estimate}
    e = 2^{\bar{M}}/\phi
\end{equation} 
where $e$ distinct cardinality and $\bar{M}$ is the mean of the registers.

In this work, we utilize a variant of Flajolet–Martin sketch; since multiple Monte-Carlo simulations are performed to calculate estimated influence, we use one register per simulation, and take the average of longest leading zero counts to calculate the average cardinality of reachability sets. Merge operations are performed if and only if there is an edge between the respective simulation.

@@

\subsection{Hash-based Fused Sampling}
The probabilistic nature of cascade models requires sampling sub-graphs $\hat{G}$ from $G = (V, E)$ to simulate the diffusion process.  Sampling can be a dominating factor in processing; sampling may demand multiple passes on the graph, if samples are memoized, multiple times of the graph size needs to be stored.
In this work, we borrow hash-based fused sampling from {\sc InFuseR}\cite{infuser}. By using a hash-based fused sampling, we eliminate the necessity of in-memory creation and storage of the sample sub-graphs.
While processing an edge, it is considered for all possible samples.
It is sampled or skipped depending on the outcome of the hash-based random value. The hash function used is given in equation \ref{eq:hash}.
\begin{equation}
    \label{eq:hash}
    h(u,v) = \mbox{{\sc Murmur3}}(u||v)~mod~2^{31}  
\end{equation}
where $||$ is the concatenation operator.

@@

For performance considerations, edge hash values were pre-computed in our work and only existing pairs were considered. The trade-off between extra memory and computation here maybe not be applicable for different computation architectures and faster/simpler hash functions.

@@

Multiple deterministic random values for each edge is generated by using the edge hash values and random numbers $X_r$ associated with each simulation $r$. 
Sampling probability of $(u, v)$ for simulation $r$, $P(u, v)_r$, is computed as follows; First, $h(u,v)$ is XOR'ed with a uniformly randomly chosen $X_r \in_R [0, h_{max}]$ and the result is normalized by dividing the value to the upper limit of the hash value $h_{max}$. Formally,
\begin{equation}
    \label{eq:hash_prob}
    P(u,v)_r = \frac{X_r \oplus h(u,v)}{h_{max}}.
\end{equation}

% The edge $\{u,v\}$ is verified to be in the sample if ${\rho}(u,v)_r$ is smaller than or equal to the threshold $w_{u,v}$. 
The edge $\{u,v\}$ is exists in the sample $r$ if and only if  ${P}(u,v)_r$ is smaller than the edge threshold $w_{u,v}$. One of the benefits of this approach is, sampling an edge is only single XOR and compare-greater-than operations. Also, the following control flow branch can be removed using conditional move operations. 
An efficient implementation of this approach will be discussed later in this section.
\begin{figure}[!ht] 
    \centering
    \includegraphics[width=1\linewidth]{./images/CDF.pdf}
    \caption{Cumulative probability function of hash-based sampling probabilities on various real-life networks.}
    \label{fig:prob-cdf} 
\end{figure}
Using a strong hash function such as {{\sc Murmur3}} ensures all bits independently change if the input is changed. This property allows us to generate good enough pseudo-random values for fair sampling. To prove the fairness of random values generated with the hash-based approach, we generated a large number of samples for various real-life networks and graphed the bias of random values $P(u,v)_r$ used while sampling.

% For a given graph $G = (V,E)$, the bias of a sampling probability $x$ is computed as $|(31/2) − E(|H(m)|) / E(|H(m)|))|$ for all $(u,v) \in E$ and $0 \leq r < R$. 
Figure~\ref{fig:prob-bias} shows the bias for 12 real-life networks. The sampling probability distribution with hash-based computation is almost identical to the uniform distribution with small bias.

@@

%Traditional implementations sample each edge in $G$ once per simulation and store them to construct a sampled subgraph. n
Being able to generate infinitely many samples on-the-fly, allows us to avoid many memory accesses. The only down-side of hash-based fused sampling is that we have to generate all random values,   $P(u,v)_r$, for each edge traversal.  Fortunately, these computations are quite fast on modern computing hardware.

\subsection{Reachability Set Cardinality Estimation}
A greedy solution to the influence maximization problem requires finding a vertex that maximizes marginal influence gain at each step until the seed set size reaches $K$. And, Single vertex influence estimation is the same problem as the count-distinct problem applied to all sample sub-graphs. So, it is possible to use Count-distinct sketches to estimate influence scores in the greedy method.

We propose an influence maximization method, that utilizes the Count-distinct sketch we mentioned in section \ref{sec:sketch} to estimate averages of distinct elements in sample subgraphs for both time and space efficiency. Algorithm \ref{algo:main} describes the method proposed as follows;

First, reachability sets of all vertices are initialized with vertices themselves. For all vertices $u$, its $j$th register is set to $M_u[j]=clz(h_j(u))$ meaning $R_{\hat{G}_j}(u) = \{u\}$. Then, we perform diffusion process on the sketch registers.

Second, the diffusion process is performed between vertex sketches. All vertices added to live vertices set $L$. Then, incoming edges $(u,v)$ for all live vertices are processed. For all simulation $u$ is live, the sketch of both sides are merged and set to $u$ registers. If any of the sketch-registers change, vertex $u$ is added to the next iteration's live vertices set $L'$. After live edges are processed, if the next iteration's live vertices set is smaller than the convergence threshold $\epsilon_c$, the diffusion process concluded. If not, we swap $L$ and $L'$, clear $L'$, and then diffusion processed is repeated.

>>>>

\IEEEtitleabstractindextext{%
\begin{abstract}
Influence maximization~(IM) is the problem of finding a seed vertex set that maximizes the expected number of vertices influenced under a given diffusion model. Due to the NP-Hardness of finding an optimal seed set, approximation methods are frequently used for IM. In addition to high-quality yet expensive approximation algorithms, to meet with the scale of today's networks, lightweight, sketch-based approaches, which do not simulate the process exactly, have been proposed in the literature. As expected, the quality of the seed sets for the latter is inferior to that of the former. In this work, we describe a fast, error-adaptive approach that leverages Count-Distinct sketches and the recently proposed hash-based fused sampling. To estimate the vertex set sizes influenced throughout the simulations, we use per-vertex Flajolet-Martin sketches where each sketch register corresponds to a sample subgraph. During the diffusion simulations, the reach set cardinalities of a single vertex are stored in adjacent registers in memory. This allows the proposed algorithm to find the number of influenced vertices in a single simulation step in different simulations at once. In addition, thanks to its efficiency and scalability of our parallel implementation, we can rebuild the sketches after observing large estimation errors. Our experimental results show that the proposed method finds high-quality seed sets while being up-to $1.7\times$--$27.9\times$ faster on average than a state-of-the-art approximation algorithm. In addition, it is $3.5\times$--$11.1\times$ faster than a sketch-based approach while producing seed sets with $3\%$--$12\%$ better influence scores.
\end{abstract}
% Note that keywords are not normally used for peer review papers.
}

@@

The study of efficient information/influence propagation in networks becomes an important area of research with several applications in many fields, such as viral marketing~\cite{leskovec2007dynamics, trusov2009effects}, social media analysis~\cite{zeng2010social, moreno2004dynamics}, and recommendation systems~\cite{lu2012recommender}.
As the study of these networks is imperative for educational, political, economic, and social purposes, a high-quality seed set to initiate the diffusion may have critical importance.
Furthermore, since the diffusion analysis may be time-critical, or increasing its coverage may be expensive in terms of computing resources and time, novel and efficient approaches to find good vertex sets that propagate the information/influence effectively are essential.

Influence maximization is the problem of finding a subset $S \subset V$ of $K$ vertices in a graph $G = (V, E)$ with the vertex set $V$ and edge set $E$ such that $S$ reaches the maximum reachability under some diffusion model $M$. In addition to introducing the problem, Kempe et al.~\cite{kempe2003maximizing} proved it to be NP-hard and provided a greedy Monte-Carlo approach that has a constant approximation for the optimal solution. This greedy approach is one of the most frequently applied algorithms for IM. The time complexity of the greedy algorithm, estimating the $\sigma$ influence score, running $R$ simulations, and selecting $K$ seed vertices is $\mathcal{O}(KRn\sigma)$ for a graph with $n$ vertices. Although they perform well in terms of seed-set quality, as a consequence, the greedy Monte-Carlo solutions are impractical for real-life networks featuring millions of vertices. Due to this reason, many heuristics and proxy methods have been proposed in the literature~\cite{MixGreedy, narayanam2010shapley, kimura2007extracting, chen2010PMIA,chen2010LDAG, kim2013scalable, cohen2014sketch, goyal2011simpath, jung2012irie,cheng2014imrank,liu2014influence,galhotra2016holistic}.

@@

In this work;
\begin{itemize}[leftmargin=*]
\item We propose \acro\footnote{\url{https://github.com/ggokturk/infuser}}, an open-source, a blazing-fast, and accurate Influence Maximization algorithm. The proposed scheme samples the edges as they are traversed across several simulations. Thus, sampling, diffusion, and count-distinct processes are fused for all simulations.

@@

That is, when the simulations traverse the same edge within close time intervals, the edge is accessed only once as long as the simulation timelines allow \acro to do so. The suggested approach, therefore, decreases the pressure on the memory management unit. In relation to the cascade model, we employ a vector compute unit, near maximum efficiency to regularize memory accesses. 
%\item In addition to instruction-level parallelism, we parallelize \acro on multi-core architectures.

\item \acro{} can process very large graphs with millions of vertices and hundreds of millions of edges under a minute, without compromising the quality of results. For large graphs, due to its race condition indifferent nature, the performance scales near linearly with the number of threads available. While processing the graph, only a few GBs of memory is used, mostly for storing the graph itself.

\item We evaluate the performance, memory consumption, and influence score with a state-of-the-art influence maximization algorithm {\sc Skim}~\cite{cohen2014sketch}, {\sc Tim+}~\cite{tim}, {\sc Imm}~\cite{minutoli2019fast} to accurately position the performance of \acro{} in the IM literature. The experiments show that \acro is $\minspeedup\times$—$\maxspeedup\times$ faster than state-of-the-art while still being better in comparison in terms of scores, and using a comparable amount of memory.

@@

Let $G = (V,E)$ be a directed graph where the $n$ vertices in $V$ represent the agents, and $m$ edges in $E$ represent the relations between the agents in $V$.
The neighborhood of a vertex $u \in V$ is denoted as $\Gamma_{G}(v) = \{v: (u,v) \in E\}$. 
A graph $G' = (V',E')$ is a sub-graph of $G$ if $V' \subseteq V$ and $E' \subseteq E$. 
Diffusion probability on the edge $(u, v) \in G$ is noted as $w_{u,v}$.
$w_{u,v}$ can be determined either by the diffusion model or according to the strength of $u$ and $v$'s relationship in the data.

\begin{table}[!ht]
    \caption{Table of notations}
    \label{tab:notation}
    \centering
    \begin{tabular}{|l|p{0.7\linewidth}|}
        \hline
        Variable & Definition  \\
        \hline
        $G = (V,E)$     & Graph $G$ with vertices $V$ and edges $E$ \\
        $\Gamma_G(v)$   & Neighborhood of incoming edges to vertex $v$ in graph $G$\\ %%FIXME
        $w_{u,v}$       & Probability of $u$ directly influencing $v$ \\
        %$SCC(v) $       & Strongly connected component of vertex $v$\\
        $R_{G}(v)$      & Reachability set of vertex $v$ on graph $G$\\
%        $\overline{R_{G}(v)} & Complement of the vertex set $R_{G}(v)$\\
        \hline\hline
        $S$             & Seed set to maximize influence\\
        $K$             & Size of the seed set\\
        $\mathcal{R}$   & Number of Monte-Carlo simulations performed\\
        $\sigma_{G}(S)$ & Influence score of $S$ in $G$, i.e., expected number of vertices reached from $S$ in $G$\\
        % $\sigma_{G}{(S,v)}$          & Marginal influence gain by adding vertex $v$ to seed set $S$\\
        \hline\hline
        $w$             & Sampling probability for edges\\
        $P(s,v)_r $     & Random probability generated for selecting edge vertices $s$ to $v$ in simulation $r$\\
        $h(u,v)$        & Hash function for edge $\{u,v\}$\\
        $h_{max}$       & Maximum value hash function $h$ can return\\
        %$X_r$           & Random number/hash generated for simulation $r$  \\
        \hline\hline
        $e$             & Estimated reachability set size\\
        % $[a, \ldots, a]_B$      & Vector of size $B$, contains all $a$\\
        $M_u[j]$        & $j$th sketch register for vertex $u$\\
        $\varsigma $    & Influence gained before last sketch build\\
        $\sigma $       & Influence Score\\
        $\delta$        & Marginal Gain after last sketch build\\
        $Err_l$         & Local estimation error of the sketch\\
        $Err_g$         & Global estimation error of the sketch\\
        $\epsilon_{g}$    & Global estimation error threshold\\
        $\epsilon_{l}$    & Local estimation error threshold\\ 
        $\epsilon_{c}$    & Non-convergenced vertex threshold\\
      %  $x[[i, j]]$   & Slice of vector  $x$ between indices $i$ and $j$, not including $j$ \\ 
        \hline         
    \end{tabular}
\end{table}
\subsection{Influence Maximization}

Influence Maximization aims to find a seed set $S \subseteq V$ among all possible size $K$ subsets of $V$ that maximizes an {\em influence spread function} $\sigma$  when the diffusion process is initiated from $S$. Although we focus on undirected graphs, for IM, the graph can be directed or undirected depending on the initial construction. Figure~\ref{fig:xx} shows a~(Fig.~\ref{fig:ic}) and directed~(Fig.~\ref{fig:wc})
graph in which the weights on the edges are diffusion/influence probabilities.

\begin{figure}[!ht] 
    \centering
  \subfloat[\small{IC}\label{fig:ic}]{%
       \includegraphics[width=0.50\linewidth]{images/ic.pdf}}
  \subfloat[\small{WC}\label{fig:wc}]{%
        \includegraphics[width=0.50\linewidth]{images/wc.pdf}}
    \\% IMAGES HERE
  \caption{\protect\subref{fig:ic} 
The directed graph $G = (V, E)$ for Independent Cascade with independent diffusion probabilities. 
\protect\subref{fig:wc}
The directed graph was obtained from the directed one by setting the diffusion probabilities of incoming edges to $1 / |\Gamma_G(v)|$ for each vertex $v \in V$. 
  }
  %\label{fig_ic_model} 
  \label{fig:xx} 
\end{figure}
In the literature, {\em independent} and {\em weighted cascade}~(IC and WC), and 
{\em linear threshold}~(LT)~\cite{kempe2003maximizing} are some of the widely recognized diffusion models for IM.

@@

\item{\bf Linear threshold} generalizes the independent cascade model and activates the $v$ vertex once the cumulative activation coming from the neighbors of $v$ exceeds the $\theta v$ threshold. 
All $\{u, v\}$~(or $(u, v)$ in the directed graph) edges 
with active $u$ vertices are taken into account in the process. Vertex $v$ is activated~\cite{kempe2003maximizing} when the number of these edge weights reaches $\theta v$.  
Therefore, the independent cascade model is a special variant of the linear threshold model with $\theta v = 0 $.

\item {\bf Triggering} proposed by Kempe et al.\cite{kempe2003maximizing} to generalize of the previous models mentioned above. In the triggering model, each neighbor has a probability to influence the vertex $v$. The diffusion process chooses a random subset of vertices called the "triggering set" to activate the vertices at each instance.
\end{itemize}

The complexity analysis stays consistent for many diffusion models including {\em Independent Cascade}~(IC), {\em Linear Threshold}~(LT),  models. We concentrate on the independent cascade model in this paper, but the proposed methods are also relevant to other models in the literature.

\subsection{Count-Distinct Sketch}\label{sec:sketch}
Count-distinct problem is the problem of finding the number of distinct elements in a stream with non-distinct elements. Computing the reachability sets of a vertex is a similar problem; the cardinality of all vertices visited is calculated while traversing sample subgraphs from the given vertex. Exact cardinality calculations require memory proportional to the cardinality.

@@

The methods based on Reverse Reachability and Bottom-Up Traversal utilize this property directly to merge the reachability sets of connected vertices to estimate vertices influence.
MixGreedy method goes one step further; it utilizes that all vertices in a connected component that have the same reachability set in undirected graphs. Therefore, for a sample graph, all reachability sets can be found in a single graph traversal.

For directed graphs, storing reachability sets for all vertices and merging these sets are infeasible for nontrivial graphs. 
If one-hot vectors are used to store reachability sets for constant insertion time, $O(n^2\mathcal{R})$ bits of memory is required and each merge operation has $O(n)$ time complexity. 
if Disjoints sets are used for storing reachability sets; $(n\bar{\sigma}\mathcal{R})$ memory is required to store all reachability sets, and each merge operation has $O(Ackermann'(\sigma))$ complexity.

@@

Flajolet–Martin algorithm only stores how rare elements are in a multiset.The rarity of the elements is often computed by counting the leading zeros in the item's hash value.  The registers are initialized with zero values in the beginning. Items are added by storing the longest leading zero count. Cardinality estimation can be done by taking 2 to the power of the stored value. Multiple registers and hash functions, $M[j]$ and $h_j$, are commonly used to reduce variance.
Adding the item $x$ to sketch $M$, and its result $M'$ is shown in equation~\ref{eq:sketch-add}.

@@

The merge operation for two sketches, $M_u$,$M_v$, can be performed by computing the pairwise maximum of the registers. Formally; 
\begin{equation}
    \label{eq:sketch-merge}
    M'[j] = max(M_u[j],M_v[j]) ~ \forall j\in(0,J]
\end{equation} 
where $M'$ is union of set and $J$ is the number of sketch registers.

Average of the longest running zeros can be used compute the cardinality and the result is divided to the correction factor $\phi \approx 0.77351$ to correct hash collisions.
\begin{equation}
    \label{eq:sketch-estimate}
    e = 2^{\bar{M}}/\phi
\end{equation} 
where $e$ distinct cardinality and $\bar{M}$ is the mean of the registers.

In this work, we utilize a variant of Flajolet–Martin sketch; since multiple Monte-Carlo simulations are performed to calculate the estimated influence, we use one register per simulation and take the average of longest leading zero counts to calculate the average cardinality of reachability sets. Merge operations are performed if and only if there is an edge between the respective simulations.

@@

\subsection{Hash-based Fused Sampling}
The probabilistic nature of cascade models requires sampling subgraphs $\hat{G}$ from $G = (V, E)$ to simulate the diffusion process. The sampling part of the algorithm can be a dominating factor in processing; sampling may demand multiple passes on the graph, if samples are memoized, multiple times of the graph size needs to be stored.
In this work, we borrow hash-based fused sampling from {\sc InFuseR}\cite{infuser}. By using a hash-based fused sampling, we eliminate the necessity of in-memory creation and storage of the sample subgraphs.
While processing an edge, it is considered for all possible samples.
It is sampled or skipped depending on the outcome of the hash-based random value. The hash function used is given in equation \ref{eq:hash}.
\begin{equation}
    \label{eq:hash}
    h(u,v) = \mbox{{\sc Murmur3}}(u||v)~mod~2^{31}  
\end{equation}
where $||$ is the concatenation operator.

@@

For performance considerations, edge hash values were precomputed in our work and only existing pairs were considered. The trade-off between extra memory and computation here maybe not be applicable for different computation architectures and faster/simpler hash functions.

@@

Multiple deterministic random values for each edge are generated by using the edge hash values and random numbers $X_r$ associated with each simulation $r$. 
Sampling probability of $(u, v)$ for simulation $r$, $P(u, v)_r$, is computed as follows; First, $h(u,v)$ is XOR'ed with a uniformly randomly chosen $X_r \in_R [0, h_{max}]$ and the result is normalized by dividing the value to the upper limit of the hash value $h_{max}$. Formally,
\begin{equation}
    \label{eq:hash_prob}
    P(u,v)_r = \frac{X_r \oplus h(u,v)}{h_{max}}.
\end{equation}

% The edge $\{u,v\}$ is verified to be in the sample if ${\rho}(u,v)_r$ is smaller than or equal to the threshold $w_{u,v}$. 
The edge $\{u,v\}$ is exists in the sample $r$ if and only if  ${P}(u,v)_r$ is smaller than the edge threshold $w_{u,v}$. One of the benefits of this approach is that an edge can be sampled using a single XOR and compare-greater-than operation. Moreover, the following control flow branch can be removed using conditional move operations. 
An efficient implementation of this approach will be discussed later in this section.
\begin{figure}[!ht] 
    \centering
    \includegraphics[width=1\linewidth]{./images/CDF.pdf}
    \caption{Cumulative probability function of hash-based sampling probabilities on various real-life networks.}
    \label{fig:prob-cdf} 
\end{figure}
Using a strong hash function such as {{\sc Murmur3}} ensures that all bits independently change if the input is changed. This property allows us to generate good enough pseudo-random values for fair sampling. To prove the fairness of random values generated with the hash-based approach, we generated a large number of samples for various real-life networks and graphed the bias of random values $P(u,v)_r$ used while sampling.

% For a given graph $G = (V,E)$, the bias of a sampling probability $x$ is computed as $|(31/2) − E(|H(m)|) / E(|H(m)|))|$ for all $(u,v) \in E$ and $0 \leq r < R$. 
Figure~\ref{fig:prob-bias} shows the bias for 12 real-life networks. The sampling probability distribution of hash-based computation is almost identical to the uniform distribution with small bias.

@@

%Traditional implementations sample each edge in $G$ once per simulation and store them to construct a sampled subgraph. n
Being able to generate infinitely many samples on-the-fly allows us to avoid many memory accesses. The only downside of hash-based fused sampling is that we have to generate all random values,   $P(u,v)_r$, for each edge traversal.  Fortunately, these computations are quite fast on modern computing hardware.

\subsection{Reachability Set Cardinality Estimation}
A greedy solution to the influence maximization problem requires finding a vertex that maximizes the marginal influence gain at each step until the seed set size reaches $K$. And, single vertex influence estimation is the same problem as the count-distinct problem applied to all sample subgraphs. So, it is possible to use Count-distinct sketches to estimate influence scores in the greedy method.

We propose an influence maximization method that utilizes the Count-distinct sketch we mentioned in section \ref{sec:sketch} to estimate the averages of distinct elements in sample subgraphs for both time and space efficiency. Algorithm \ref{algo:main} describes the method proposed as follows;

First, the reachability sets of all vertices are initialized with the vertices themselves. For all vertices $u$, its $j$th register is set to $M_u[j]=clz(h_j(u))$ meaning $R_{\hat{G}_j}(u) = \{u\}$. Then, we perform the diffusion process on the sketch registers.

Second, the diffusion process is performed between vertex sketches. All vertices added to live vertices set $L$. Then, the incoming edges $(u,v)$ for all live vertices are processed. For all simulations where $u$ is live, the sketch of both sides are merged and set to $u$ registers. If any of the sketch registers change, the vertex $u$ is added to the next iteration's live vertices set $L'$. After the live edges are processed, if the next iteration's live vertices set is smaller than the convergence threshold $\epsilon_c$, the diffusion process concluded. If not, we swap $L$ and $L'$, clear $L'$, and then the diffusion processed is repeated.

In Greedy algorithm\cite{kempe2003maximizing} for each candidate vertex, influence diffusion computations are performed for each vertex.  On the other hand, diffusion on sketch registers allows finding all vertices reachability set. All vertices influence scores in a single-step distance only computed once. This dynamic programming like property allows fast selection of influence maximizing candidates. Only iterations as much as the diameter of sample graphs are required to converge the sketches.
